# Estimating and reporting

## Generating some fake data

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}


#maybe replace with
# https://egap.github.io/learningdays-resources/Slides/estimation-slides.html#/using-declaredesign-define-estimand-and-estimators
library(tidyverse)

rm(list=ls())
set.seed(1)

data <-  
  read_csv("data/SAFI_clean.csv", na = "NULL") %>%
  left_join({.} %>%
            select(village) %>%
            distinct(village) %>%
            rowwise %>%
            mutate(treatment = rbinom(1,1,0.5)))%>%
  rowwise() %>%
  mutate(educated = rbinom(1,1,0.3),
         u = rnorm(1,sd=0.5),
         prob = 0.3 * treatment + 0.1 * educated + u,
         prob = max(0,prob),
         prob = min(1,prob),
         uses_fertilizer = rbinom(1,1,prob))        ) 

```


## Making a table of summary statistics

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}


library(gtsummary)

data %>%
  select(where(is.numeric)) %>%
  tbl_summary( statistic = list(all_continuous() ~ "{mean} ({sd})"))


data %>%
sumtable(out = "return") %>%
huxtable() %>%
as_flextable() %>%
save_as_docx(path = "assets/summ_table.docx")


library(table1)


table1(data = data) 


library(tableone)

data%>%
select(where(is.numeric)) %>%
CreateTableOne(data=.,strata="treatment") %>%
print(printToggle = FALSE) %>%
as.data.frame() %>% 
rownames_to_column("Variables")  %>% 
flextable::as_flextable() %>%
save_as_docx(path = "assets/summ_table.docx")


library(furniture)

data%>%
select(where(is.numeric)) %>%
furniture::table1(data)

```

## Simple regression

A simple regression uses the `lm()` function.

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}

lm <- lm(uses_fertilizer ~ treatment + educated, data = data)
summary(lm)


```

## Robust standard errors

To get robust standard errors clustered at the village
level, we can use the following:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}

library(lmtest)
library(sandwich)

coeftest <- coeftest(fit, vcov = vcovHC(fit, type = 'HC0'),cluster=~village)
coeftest
```

Or, alternatively, we can use the `lm_robust()` from the 
DeclareDesign library:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}

library(DeclareDesign)

lm_robust <- lm_robust(uses_fertilizer ~ treatment + educated, 
                        cluster = village,
                        data = data)

summary(lm_robust)

```
## Exporting to word

To export the resulting tables to word, I use the huxtable and
flextable packages.

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}

library(flextable)
library(huxtable)

huxreg(lm,coeftest,lm_robust,
       statistics = c(N = "nobs", "Adj. R2" = "adj.r.squared")) %>%
as_flextable() %>%
save_as_docx(path = "assets/regression_table.docx")

```






