# Estimating and reporting {#reporting}

This chapter uses a large number of packages, and the SAFI data 
set, so make sure
all are downloaded by running the code from the [Set-up section](#setup).

I will create a table of descriptive statistics, and a simple regression
table.

## Generating some fake data

First we make a fake intervention aimed at improving fertilizer adoption.
Adoption depends on the treatment and education and a random component.
The page on [DeclareDesign](#declaredesign) has more advanced techniques
to generate fake data.

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}

library(tidyverse)
library(here)

rm(list=ls())
set.seed(1)

data <-  
  read_csv(here("data/SAFI_clean.csv"), na = "NULL") %>%
  left_join({.} %>%
            select(village) %>%
            distinct(village) %>%
            rowwise %>%
            mutate(treatment = rbinom(1,1,0.5)))%>%
  rowwise() %>%
  mutate(educated = rbinom(1,1,0.3),
         u = sample(c(0.1,0.2,0.3),1),
         prob = 0.3 * treatment + 0.1 * educated + u,
         uses_fertilizer = rbinom(1,1,prob)) %>%
  ungroup() %>%
  select(-prob,-u) 

```

## Making a table of summary statistics


### Using Arsenal
The first table we make is a table of descriptive statistics
using the Arsenal package:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=FALSE}

library(arsenal)

descriptive_table <-
  data %>%
  select(where(is.numeric),-key_ID) %>%
  tableby(treatment ~ ., 
          data = .)

summary(descriptive_table)

```

That summary looks okay, even as raw text, and does what we want, so
let's export it to word! For this, we use Arsenal's `write2word()`
function. I had some trouble getting it to pick up my path, so I 
am using the [here library](https://here.r-lib.org/) to generate 
a path for me. This library is a must for shared projects!


```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=FALSE}

write2word(descriptive_table, here("tables/summ_stats.docx"), title = "Descriptive Statistics", quiet = TRUE)

```

You can find the resulting .docx [here](https://github.com/kleuveld/r_cheatsheet/raw/main/tables/summ_stats.docx).


### Using Flextable

For more control, the flextable package can covert data frames into good-looking table using the 
`tabulator()` function.

First, make a data frame with summarry statistics. I duplicate the data set using `bind_rows()`
to create an overall group.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

summstats <-
bind_rows(data %>% mutate(Treatment = ifelse(treatment," Treatment","  Control")),
          data %>% mutate(Treatment = "Overall")) %>% 
  select(where(is.numeric),Treatment,-key_ID,-treatment) %>%
  group_by(Treatment) %>%
  summarize(across(everything(),
                     list(n =  ~sum(!is.na(.x)),
                          mean = ~mean(.x,na.rm=TRUE),
                          sd = ~sd(.x,na.rm=TRUE)),
                      .names =  "{.col}-{.fn}")) %>%
  pivot_longer(cols = -Treatment,
                 names_to = c("Variable",".value"),
                 names_sep="-") 



```


Then use flextable. Note that `tabulator()` sorts the columns alphabetically, so that would be
control, overall, treatment. That doesn't make sense, so I have used spaces (`" Treatment"`) to 
control the ordering. I also haven't found a way to find significant differences.


```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

library(flextable)
summstats %>%
tabulator(rows = "Variable",
                    columns = "Treatment",
                    `N` = as_paragraph(as_chunk(n,digits=0)),
                    `Mean (SD)` =  as_paragraph(as_chunk(fmt_avg_dev(mean, sd, digit2 = 2)))) %>%
as_flextable() 


```

## Simple regression

A simple regression uses the `lm()` function. We save the results
in an object, which we can later include in a table we export to word.

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=FALSE}

lm <- lm(uses_fertilizer ~ treatment + educated, data = data)
summary(lm)


```

## Robust standard errors

To get robust standard errors (bootstrapped) clustered at the village
level, we can use the following:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=FALSE}

library(lmtest)
library(sandwich)

lm_robust <- coeftest(lm, vcov = vcovBS(lm, cluster=~village))
lm_robust

```
<!-- alternatively: use lm_robust from estimatr -->


Let's also do a probit model:


```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}

probit <- glm(uses_fertilizer ~ treatment + educated, 
                family = binomial(link = "probit"), 
                data = data)

summary(probit)

```

And get robust SEs:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}


probit_robust <- coeftest(probit, vcov = vcovBS(probit,cluster=~village))
probit_robust

```


## Exporting to word

To make a good looking regression table, I use the huxtable 
library. Note the use of `tidy_override()` to add statistics
to the model output.

To export the resulting tables to word, I use the huxtable package.
The package is extremely flexible, so check out [its website](https://cran.r-project.org/web/packages/huxtable/vignettes/huxreg.html).
Some of the flexibility is seen below, by adding the number of clusters
and formatting the number in the resulting table.

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=FALSE}

#add the number of clusers to each
lm_robust <- 
  huxtable::tidy_override(lm_robust,
                glance = list(n_clusters = 4),
                extend=TRUE)
probit_robust <- 
  huxtable::tidy_override(probit_robust,
                glance = list(n_clusters = 4),
                extend=TRUE)
#make the table
reg_table <-
  huxtable::huxreg("Linear" = lm,"Linear Robust" = lm_robust,
         "Probit" = probit, "Probit Robust" = probit_robust,
         statistics = c(N = "nobs",
                        Clusters = "n_clusters", 
                        "Adj. R2" = "adj.r.squared")) %>% 
  huxtable::set_number_format(row=9,value=0)

reg_table

```

That looks decent! However, huxtable doesn't work well with
Word. We can use `as_flextable()` to convert to flextable
(note that I had bad results with `flextable::as_flextable()`, so make
sure to use `huxtable::as_flextable()`), and then use flextable
to export to docx. 

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=FALSE}

reg_table %>%
  huxtable::as_flextable() %>%
  flextable::save_as_docx(path = here("tables/regression_table.docx"))

```

You can find there output [here](https://github.com/kleuveld/r_cheatsheet/raw/main/tables/regression_table.docx).