# Plotting using ggplot {#plotting}

## Introduction

Each chart built with ggplot2 must include the following:

    - Data
    - Aesthetic mapping (aes)
    - Geometric objects (geom)

Thus, the template for graphic in ggplot2 is:

```

<DATA> %>%
    ggplot(aes(<MAPPINGS>)) +
    <GEOM_FUNCTION>()

```

## Data Prep

```{r}

library(tidyverse)
library(here)

data <- 
  read_csv(here("data/SAFI_clean.csv"), na = "NULL") %>%
    separate_longer_delim(items_owned, delim = ";") %>%
    mutate(value = 1) %>%
    pivot_wider(names_from = items_owned,
                values_from = value,
                names_glue = "owns_{items_owned}",
                values_fill = 0) %>%
    rowwise %>%
    select(-"owns_NA") %>%
    mutate(number_items = sum(c_across(starts_with("owns_"))))  


```


## Scatter plot

A scatter plot uses `geom_point()`

```{r}


data %>%
    ggplot(aes(x = no_membrs, y = number_items)) +
    geom_point()


```

To differentiate overlapping points, you can use transparency...

```{r}

data %>%
    ggplot(aes(x = no_membrs, y = number_items)) +
    geom_point(alpha = 0.3)

```

...or add jitter:

```{r}

data %>%
    ggplot(aes(x = no_membrs, y = number_items)) +
        geom_jitter()

```

To add groups, add the variable that defines the groups as an aesthetic mapping, either
in the call to ggplot(), or the one in the `geom_()` function you use:

```{r}

data %>%
    ggplot(aes(x = no_membrs, y = number_items)) +
    geom_jitter(aes(color = village))

```

## Boxplot

```{r}

data %>%
    ggplot(aes(x = respondent_wall_type, y = rooms)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3,
                color = "tomato",
                width = 0.2,
                height = 0.2)


```

## Bar chart

For a simple bar chart of counts:

```{r}

data %>%
    ggplot(aes(x = respondent_wall_type)) +
    geom_bar(aes(fill = village))

```

And since stacked bar charts are not easy to read:

```{r}

data %>%
    ggplot(aes(x = respondent_wall_type)) +
    geom_bar(aes(fill = village), position = "dodge")

```

Note that `geom_bar()` defaults to displaying counts. If you want something else, you can
use the `stat =` option. `stat = "identity"` is especially useful, as it displays values as-is,
allowing you to pre-process your data anyway you want, for example to get percentages: 

```{r}

wall_plot <-
  data %>%
  filter(respondent_wall_type != "cement") %>%
  group_by(village, respondent_wall_type) %>%
  summarize(n = n()) %>%
  mutate(percent = (n / sum(n)) * 100) %>%
  ungroup() %>%
  ggplot(aes(x = village, y = percent, fill = respondent_wall_type)) +
    geom_bar(stat = "identity", position = "dodge")    

wall_plot

```

## Labels

The `labs()` function changes labels, the `scale_` family of functions change axes and legends:

```{r}

wall_plot +
    labs(title = "Proportion of wall type by village",
         fill = "Type of Wall in Home",
         x = "Village",
         y = "Percent") +
    scale_fill_discrete(labels=c('Burned briks', 'Mud Daub', 'Sun Bricks')) +
    scale_y_continuous(limits= c(0,100), n.breaks = 6)

```

## Faceting {#faceting}

Faceting allows splitting a graph in multiple parts:

```{r}

data %>%
    group_by(village) %>%
    summarize(across(.cols = starts_with("owns_"),
                     .fns = ~sum(.x,na.rm=TRUE) / n() * 100,
                     .names = "{str_replace(.col, 'owns_', '')}")) %>%
    pivot_longer(-village, names_to = "items", values_to = "percent") %>%
    ggplot(aes(x = village, y = percent)) +
        geom_bar(stat = "identity", position = "dodge") +
        facet_wrap(~ items) +
        theme_bw() +
        theme(panel.grid = element_blank())


```

Note that the `.names` argument to `summarize(across())` is specified as a 
[glue string](https://glue.tidyverse.org/) that uses `str_replace()` to cut off the `"owns_"` 
bit of the column names.

## Stacked bar chart with WUR template and observation counts {#stacked-wur}

First, download the WUR template from [here](https://git.wur.nl/wmrkdown/ggthemewur). 
Install it following the instruction on that page.

Then we make a stacked bar chart using `position = position_fill()`. 
I use `reverse = TRUE` because I think the ordering doesn't make 
sense in these horizontal plots.


```{r}
library(ggthemewur)

data %>%
ggplot(aes(y = village, fill = respondent_wall_type)) +
  geom_bar(position = position_fill(reverse = TRUE))  +
  theme_wur() +
  scale_fill_wur_discrete() +
  scale_x_continuous(labels = scales::percent_format())




```

I like the percentages, but perhaps it's good to know how many observations we have in each village. We can do this by changing the village names:


```{r}

data %>%
  group_by(village) %>%
  add_count() %>%
  mutate(village = paste0(village," (n =",n,")")) %>%
  ggplot(aes(y = village, fill = respondent_wall_type)) +
  geom_bar(position = position_fill(reverse = TRUE))  +
  theme_wur() +
  scale_fill_wur_discrete() +
  scale_x_continuous(labels = scales::percent_format())

```


You can also put the counts inside the plot:

```{r}

data %>%
  ggplot(aes(y = village, fill = respondent_wall_type)) +
  geom_bar(position = position_fill(reverse = TRUE))  +
  theme_wur() +
  scale_fill_wur_discrete() +
  scale_x_continuous(labels = scales::percent_format()) +
  geom_text(stat = "count", 
            aes(label = after_stat(count)), 
            position = position_fill(vjust = 0.5, reverse = TRUE), 
            color = "white") 

```

Or as percentages:



```{r}

data %>%
  ggplot(aes(y = village, fill = respondent_wall_type)) +
  geom_bar(position = position_fill(reverse = TRUE))  +
  theme_wur() +
  scale_fill_wur_discrete() +
  scale_x_continuous(labels = scales::percent_format()) +
  geom_text(aes(label = after_stat(scales::percent(count / sum(count),accuracy=1))),
            stat = "count",
            position = position_fill(vjust = 0.5, reverse = TRUE),
            color = "white") 

```

Use [functions](#plot-function) to make all this code easier to use.


## Ordering of labels

By default, `ggplot` will order categorical variables in your graph alphatically. 
With the wall types this was fine, but when the categories have an order, this 
doesn't look good:

```{r}

data %>%
    ggplot(aes(y = village, fill = affect_conflicts)) +
        geom_bar(position = position_fill(reverse = TRUE))


```  

Here you'd expect the `once` category to be between `more_once` and `never`, not at the end.
To fix this, convert the variable to a [factor](https://myslu.stlawu.edu/~iramler/stat234/coursenotes/working-with-factors-in-r-an-introduction-to-the-forcats-package.html).
The order in the `levels` argument will be the order in which the labels will be displayed:

```{r}

data %>%
    mutate(affect_conflicts = factor(affect_conflicts, levels = c("frequently",
                                                                  "more_once",
                                                                  "once",
                                                                  "never"))) %>%
    ggplot(aes(y = village, fill = affect_conflicts)) +
        geom_bar(position = position_fill(reverse = TRUE))  
   

```

Note that the trick of adding group counts to string variables using  `add_count()` doesn't work with factors.
Or rather, it works, but converts the factor back to a string. 
In the [functions](#plot-function) chapter, I define a function called `count_label()` that adds group counts 
to factor variables using black magic. 

## Maps

For maps, we use the `sf` package, and a sample data set, in geo-json format (but sf can use
all sorts of shapefiles).

```{r}

library(sf)

file <- "https://raw.githubusercontent.com/johan/world.geo.json/master/countries.geo.json"


shapefile <- st_read(file)

shapefile

```

You can use `ggplot()` and `geom_sf()` to make a map:

```{r}

shapefile %>%
    ggplot() +
    geom_sf()

```

You can use the shapefile as a regular data file, using any old data wrangling 
functions on it. 


```{r}

shapefile_updated <-
    shapefile %>%
    mutate(x = rnorm(n = nrow(.))) 

shapefile_updated

```
You can use the `fill` aesthetic to color your shapefile:

```{r}

shapefile_updated %>%
    ggplot(aes(fill = x)) +
        geom_sf(colour = NA) +  # removes borders
        theme_void()            # removes grid

```

You can also make an interactive map, which you can use in html documents created with rmarkdown,
or in shiny applications. It uses a palette I created using the `colorBin()` function.

```{r}

library(leaflet)


pallete <- colorBin(
  palette = "YlOrBr", domain = shapefile_updated$x,
  na.color = "transparent", bins = 5
)

shapefile %>%
    mutate(x = rnorm(n = nrow(.))) %>%
    leaflet() %>%
    addTiles() %>%
    addPolygons(fillColor = ~pallete(x),
                stroke = TRUE,
                fillOpacity = 0.9,
                color = "white",
                weight = 0.3) %>%
    addLegend(pal = pallete, values = ~x, opacity = 0.9,
              title = "Look at these pretty colours", position = "bottomleft")


```