# Plotting using ggplot {#plotting}

## Introduction

Each chart built with ggplot2 must include the following:

    - Data
    - Aesthetic mapping (aes)
    - Geometric objects (geom)

Thus, the template for graphic in ggplot2 is:

```

<DATA> %>%
    ggplot(aes(<MAPPINGS>)) +
    <GEOM_FUNCTION>()

```

## Data Prep

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

library(tidyverse)
library(here)

data <- 
  read_csv(here("data/SAFI_clean.csv"), na = "NULL") %>%
    separate_longer_delim(items_owned, delim = ";") %>%
    mutate(value = 1) %>%
    pivot_wider(names_from = items_owned,
                values_from = value,
                names_glue = "owns_{items_owned}",
                values_fill = 0) %>%
    rowwise %>%
    select(-"owns_NA") %>%
    mutate(number_items = sum(c_across(starts_with("owns_"))))  


```


## Scatter plot

A scatter plot uses `geom_point()`

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}


data %>%
    ggplot(aes(x = no_membrs, y = number_items)) +
    geom_point()


```

To differentiate overlapping points, you can use transparency...

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

data %>%
    ggplot(aes(x = no_membrs, y = number_items)) +
    geom_point(alpha = 0.3)

```

...or add jitter:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

data %>%
    ggplot(aes(x = no_membrs, y = number_items)) +
        geom_jitter()

```

To add groups, add the variable that defines the groups as an aesthetic mapping, either
in the call to ggplot(), or the one in the `geom_()` function you use:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

data %>%
    ggplot(aes(x = no_membrs, y = number_items)) +
    geom_jitter(aes(color = village))

```

## Boxplot

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

data %>%
    ggplot(aes(x = respondent_wall_type, y = rooms)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3,
                color = "tomato",
                width = 0.2,
                height = 0.2)


```

## Bar chart

For a simple bar chart of counts:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

data %>%
    ggplot(aes(x = respondent_wall_type)) +
    geom_bar(aes(fill = village))

```

And since stacked bar charts are not easy to read:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

data %>%
    ggplot(aes(x = respondent_wall_type)) +
    geom_bar(aes(fill = village), position = "dodge")

```

Note that `geom_bar()` defaults to displaying counts. If you want something else, you can
use the `stat =` option. `stat = "identity"` is especially useful, as it displays values as-is,
allowing you to pre-process your data anyway you want, for example to get percentages: 

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

wall_plot <-
  data %>%
  filter(respondent_wall_type != "cement") %>%
  group_by(village, respondent_wall_type) %>%
  summarize(n = n()) %>%
  mutate(percent = (n / sum(n)) * 100) %>%
  ungroup() %>%
  ggplot(aes(x = village, y = percent, fill = respondent_wall_type)) +
    geom_bar(stat = "identity", position = "dodge")    

wall_plot

```

## Labels

The `labs()` function changes labels, the `scale_` family of functions change axes and legends:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

wall_plot +
    labs(title = "Proportion of wall type by village",
         fill = "Type of Wall in Home",
         x = "Village",
         y = "Percent") +
    scale_fill_discrete(labels=c('Burned briks', 'Mud Daub', 'Sun Bricks')) +
    scale_y_continuous(limits= c(0,100), n.breaks = 6)

```

## Faceting

Faceting allows splitting a graph in multiple parts:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

data %>%
    group_by(village) %>%
    summarize(across(starts_with("owns_"),~sum(.x,na.rm=TRUE) / n() * 100)) %>%
    pivot_longer(-village, names_to = "items", values_to = "percent") %>%
    mutate(items = str_replace(items,"owns_","")) %>%
    ggplot(aes(x = village, y = percent)) +
        geom_bar(stat = "identity", position = "dodge") +
        facet_wrap(~ items) +
        theme_bw() +
        theme(panel.grid = element_blank())


```