# Power Analysis using DeclareDesign

[DeclareDesign](https://declaredesign.org/) is a system to simulate Research Designs. This is useful for power 
analysis, because it is often hard to include things like clustering and 
covariates in standard power calculators. 

Resources for learning about DeclareDesign:

- [Slides by the authors of DeclareDesign: Graeme Blair, Alex Coppock, Macartan Humphreys](https://macartan.github.io/slides/202211_declaredesign_and_power.html)
- [The DeclareDesign CheatSheet](https://raw.githubusercontent.com/rstudio/cheatsheets/master/declaredesign.pdf)
-  The book [Research Design in the Social Sciences: Declaration, Diagnosis, and Redesign](https://book.declaredesign.org/)

Read at least the slides before going forward!

## Diff-in-Diff

Here's a common situation: I've collected some baseline data, and are 
wondering if our study has sufficient power to pick up expected treatment
effects.

I will start off with some fake baseline data, which only contains `y0`, our outcome
indicator. We use the `fabricate()` function, from the fabricatr library, which is
loaded with DeclareDesign:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

# start with a clean environment
rm(list=ls())

# load packages
library(tidyverse) # data management

library(fixest) # for estimation of fixed effects
library(lmtest) # for computing standard errors
library(sandwich) # provides the functions for lmtest to compute robust SEs

library(DeclareDesign) # for power calculation
library(broom) # to extract coefficients from model output

# set seed for reproducibility
set.seed(1)

N = 100
fake_data <-
  fabricate(N = N, 
            y0 = runif(N, 100, 150))

fake_data %>% as_tibble()

```

Now, let's add an extra year to our data, again using `fabricate()`, just
to demonstrate how, using `add_level()` and  then `crosslevels()`:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

fakedata_years <-
  fabricate(fake_data,
            years = add_level(N = 2, 
                              t = as.numeric(years) - 1, 
                              nest=FALSE),
            observations = cross_levels(by = join_using(ID, years)))

fakedata_years %>% as_tibble()

```

Now it's time to start declaring our design. The first element of the 
design is the model, which essentially is my data, and I declare it 
using the same synytax as `fabricate()`
above. This means you can input existing data, but also generate 
random new variables.

The potential outcomes are generated using `potential_outcomes()`.
This will create two variables `Y_Z_1` and `Y_Z_0`, which are 
the potential outcomes if having received treatment (`Z == 1`) or not. 
Note that I will generate `Z` in the next step.

You can see I have created a very simple data generating 
process to generate different outcomes in `t = 1` for treatment 
and control, but you can go absolutely wild here to check the 
assumptions you have: you can cluster the outcomes, include 
compliance or treatment effectiveness, add more or less noise, 
make it depend on  both observable and unobservable 
characteristics, etc. etc.


```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

year_shock <- 10
effect_size <- 15
sd <- 5

model <- 
  declare_model(fakedata_years,
                potential_outcomes(Y ~ y0 + t * year_shock + 
                                            Z * effect_size + 
                                            t * rnorm(N,sd = sd)))

```

Then it's time to think about assignment. I need to assign treatment, 
and reveal the outcomes based on treatment assignment. Here I create
two variables for treatment so I can toy around with different model
specifications later on. The first variable is the `treatment_group`, to 
indicate which households get the treatment. For this I use the 
`cluster_ra()` because I have two observations per household, 
so households can be thought of as a cluster. The other is `Z`, 
an indicator for actually having
received the treatment (equivalent to `Z = treatment_group * t`). 

If treatment had already been known during baseline (not
unlikely) then we would have only had to reveal the outcomes:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

assignment <-
  declare_assignment(treatment_group = cluster_ra(clusters = ID, prob = 0.5),
                     Z = ifelse(treatment_group == 1 & t == 1,1,0),
                     Y = reveal_outcomes(Y ~ Z)) 

```

Next, I declare my theoretical quantity of interest is the treatment 
effect in year 1. This is useful to check for bias. This is not really
needed here, since my model will follow our data generation process exactly,
so I could have just skipped this step.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

inquiry <-  
  declare_inquiry(ATE = mean(Y_Z_1 - Y_Z_0), subset = (t == 1)) 

```

And finally, I add three estimators. If I hadn't declared an inquiry, 
I could have declared a test here using `declare_test()`, which has 
the same syntax.

The first estimator uses `lm_robust()`, which is included
with DeclareDesign, to estimate a simple linear model using standard 
DiD notation.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

estimator1 <-
  declare_estimator(Y ~ t * treatment_group, 
                    inquiry = "ATE", 
                    .method = estimatr::lm_robust, 
                    term = "t:treatment_group", 
                    label = "LM")

```
Then I use `feols()` from the fixest package to estimate a 
Two-way fixed effects (2FE) model. 
This is equivalent to the model above for this case, but the 
2FE can be more easily extended to more complex 
designs ([but beware!](https://doi.org/10.1017/pan.2020.33)).

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

library(fixest)
library(broom)

estimator2 <-
  declare_estimator(Y ~ Z | treatment_group + t,
                    vcov = "hetero",
                    .method = feols, 
                    term = "Z", 
                    inquiry = "ATE", 
                    label = "2FE")

```

As third estimator, I use a custom estimation function. The function
takes the data to be analyzed as its `data` argument, and as its output
it returns a tidy dataset created by `broom::tidy()`. As long as you make 
sure the data can go into your custom function,
and it returns something that looks like what [`broom::tidy()`](https://broom.tidymodels.org/articles/broom.html) would
return, there's no limit as to how complex we can make your analysis.

Here, I keep it (relatively) simple. I use the lmtest/sandwich way of estimating 
robust standard errors, and make sure you can easily change the equation and type of
Standard Errors to use through a parameter.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

library(lmtest)
library(sandwich)

custom_robust <- function(equation, data, type="HC1"){
  # function takes a formula, provided as the first argument
  # of declare_estimator...
  # .. a type of standard errors to pass to vcov (default is HC1)
  # ... and data
  # It outputs a tidy data frame with relevant coefficients.

  lm(equation, data = data) %>%
    coeftest(., vcov = vcovHC(., type=type)) %>% 
    broom::tidy(conf.int=TRUE) %>%
    mutate(outcome = "Y") %>%
    filter(term == "t:treatment_group")
}

# the arguments formula and type are passed to to custom_robust()
# as is the data on which to run the estimator. 
estimator3 <- 
  declare_estimator(equation = Y ~ t * treatment_group, 
                    handler = label_estimator(custom_robust), 
                    inquiry = "ATE", 
                    label = "Custom") 

```

Finally, I combine all these elements to declare my design. Note that 
it's only here that R starts actually running the code to randomize 
things. The previous was just declaration!

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}
  
design <- model + assignment + inquiry + estimator1 + estimator2 + estimator3
summary(design)

```

Note that the results of my three estimators are identical, which is as expected.

If you want to browse a version of the data created by your design,
use the `draw_data()` function. This is useful to examine the properties of the data.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

draw_data(design) %>%
  as_tibble()

```

Now to calculate our power. The `diagnose_design()` will run our
model 500 times and our power is simply the fraction of times we
find a statistically significant effect.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

diagnose_design(design)

```

Looks good! (Except for some NAs for the 2FE model: this is due
to the summary function not providing all things needed. 
To fix this, I could have created my own summary
or handler function, but it doesn't really affect anything,
so this is good enough.)

But what if I was too optimistic? Using the 
`redesign()` function, you can vary various 
parameters of your design, and test all their combinations.
Here I check a few plausible values for 
the `sd` variable (giving the variation of the error term for 
incomes in year 1) and the expected effect size.

Note that by default, this runs all estimators, 500 times for each
combination of the parameters. This takes a lot of time, so I update
my design to have only one estimator (estimator1, which is fast) 
and I've set the `sims` option of `diagnose_design()` to 200.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

design <- model + assignment + inquiry + estimator1

diagnosis <-
  design %>%
  redesign(effect_size = 6:15,
           sd = c(3,5,7)) %>%
  diagnose_design(sims = 200)

```

Now, I want a nice plot. For this I use the
`tidy()` function to convert the output to a data frame, which -
after some filtering - I pipe into `ggplot()`. 
It's now easy to see that I should 
be able to pick up an effect size of 10 in most cases.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

diagnosis %>%
  tidy() %>%
  filter(diagnosand == "power") %>%
  select(effect_size,sd,power = estimate) %>%
  mutate(sd = factor(sd)) %>%
  ggplot(aes(x = effect_size, y = power, 
             shape = sd, color=sd)) +
    geom_line() +
    geom_point() +
    geom_hline(yintercept=0.8,linetype=2) + 
    scale_x_continuous(breaks=seq(6,16,2))

```


