# Functions

Functions are great to re-use and structure your code.

## Basic structure

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}


add_two <- function(input) {
	# Do something with the input
	output <- input + 2

	#call the intended output in the last line
	output
}

add_two(84)

```

## Using functions in a pipe

By default, a function in a pipe uses the dataset of the pipe as its first argument.
Also make sure to include a quick description of what your function does.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}
library(tidyverse)

add_number_column <- function(data,number=2) {
	#This functions takes a data frame, and returns the same data frame,
	#but with an exta column of 2sas its first variable. 
	#The user can specify a number other than 2 using the num argument
	#This is great if you want a column of twos as the first variable
	#of your data set.
	data %>%
		mutate(bestvar = number) %>%
		select(bestvar,everything())
}

read_csv("data/SAFI_clean.csv",na = "NULL") %>%
add_number_column() 

read_csv("data/SAFI_clean.csv",na = "NULL") %>%
add_number_column(4) 


```




## Passing variables to a function

The following doesn't work, because R can't find `no_membrs`, even though it 
does exist in the data set in our pipe.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=TRUE}

library(tidyverse)

add_two_to_a_column_broken <- function(data,variable) {
	#function that wants to add two two a column specified by the user
	#but it doesn't work.
	data %>%
	mutate(variable = variable + 2)
}

read_csv("data/SAFI_clean.csv",na = "NULL") %>%
add_two_to_a_column_broken(no_membrs) 


```

For a technical note on why this is, 
see [here](https://dplyr.tidyverse.org/articles/programming.html).
To solve this, wrapp the variable in your function definition
in `{{` and `}}`, and change the `=` in `mutatate` to `:=`. 


```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

library(tidyverse)


add_two_to_a_column_fixed <- function(data,variable) {
	#function that adds two two a column specified by the user
	data %>%
	mutate( {{variable}} := {{variable}} + 2)
}

read_csv("data/SAFI_clean.csv",na = "NULL") %>%
	add_two_to_a_column_fixed(no_membrs)

```


Remember our code to generate dummies based on text:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}

library(tidyverse)

read_csv("data/SAFI_clean.csv", na = "NULL") %>%
    mutate(items_owned = ifelse(is.na(items_owned),"None",items_owned)) %>%
    select(key_ID,items_owned) %>%
    separate_longer_delim(items_owned, delim = ";") %>%
    mutate(value = 1) %>%
    pivot_wider(names_from = items_owned,
                values_from = value,
                names_prefix = "owns_",
                values_fill = 0) 

```

We now want to apply this same principle to `months_lack_food` column.
We could copy-paste the code, but the better solution is to make a function:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}


create_dummies <- function(df,stringvar,prefix="",delim = ";") {
	#Function that generates dummies from a string variable
	#containing multi-response answers
	#returns dataset without original var, but with dummies.

	df %>%
		separate_longer_delim({{stringvar}}, delim = delim) %>%
	    mutate(value = 1) %>%
        pivot_wider(names_from = {{stringvar}},
                    values_from = value,
                    names_prefix = prefix,
                    values_fill = 0) 
}

read_csv("data/SAFI_clean.csv", na = "NULL") %>%
	mutate(items_owned = ifelse(is.na(items_owned),"None",items_owned)) %>%
	create_dummies(stringvar = items_owned, prefix = "owns_") %>%
	create_dummies(months_lack_food,"lack_food_")

```

The code is now clear, I can update the function in one place and all variables will
be updated accordingly. Plus, if I want to use the same code in another project, I can
easily do it. 