```{r, include = FALSE}


# start with a clean environment
rm(list=ls())

# run diagnosis?
run_diagnosis = FALSE


```


# Power Analysis using DeclareDesign {#declaredesign}

[DeclareDesign](https://declaredesign.org/) is a system to simulate Research Designs. This is useful for power 
analysis, because it is often hard to include things like clustering and 
covariates in standard power calculators. 

Resources for learning about DeclareDesign:

- [Slides by the authors of DeclareDesign: Graeme Blair, Alex Coppock, Macartan Humphreys](https://macartan.github.io/slides/202211_declaredesign_and_power.html)
- [The DeclareDesign CheatSheet](https://raw.githubusercontent.com/rstudio/cheatsheets/master/declaredesign.pdf)
-  The book [Research Design in the Social Sciences: Declaration, Diagnosis, and Redesign](https://book.declaredesign.org/)

Read at least the slides before going forward!

## Simple Power Calculation

Let's first do some power calcuations. 
I will use Rhomis data from Burundi, you can get the data 
[here](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/9M6EHS). 


### Loading data

I do some data wrangling, like creating village codes based on y-coordinates,
and some assorted cleaning.

```{r}

library(tidyverse)
library(DeclareDesign)

n_clusters = 18

rhomis <-
  read_csv(here("data/RHoMIS_Indicators.csv")) %>% #names()
  filter(Country == "Burundi") %>% 

  # split the sample in villages, by lattitude
  filter(!is.na(GPS_LAT)) %>%
  arrange(GPS_LAT) %>%
  mutate(village = rep(1:n_clusters, each = nrow(.) / n_clusters , length.out =  nrow(.))) %>%
  select(village,HHsizeMAE,HouseholdType,Head_EducationLevel,LandOwned,HFIAS_status,NFertInput, total_income_USD_PPP_pHH_Yr) %>%
  
  # create some variables
  mutate(HFIAS_status = factor(HFIAS_status, levels = c("FoodSecure", "MildlyFI", "ModeratelyFI", "SeverelyFI")),
         food_secure = 1 * (as.numeric(HFIAS_status) <= 3),
         educated = 1*(Head_EducationLevel != "No_school"),
         female = 1* (HouseholdType == "woman_single")) %>%
  mutate(hh = row_number()) %>%
  select(!HFIAS_status) %>%

  # there are too many missings in my data; these seem reasonable to assume to be 0 when missing:
  mutate(across(c(LandOwned,NFertInput,educated),
              ~ if_else(is.na(.x),0,.x)))


```

### Declare Design 

Now it's time to start declaring our design. The first element of the 
design is the model, using `declare_model()` which essentially is my data.
`declare_model()` follows the syntax of `fabricate()`, 
which allows you to generate an entirely new dataset. 
For now we stick with existing data,
but the syntax is still the same as `fabricate()`, 
so insead of using `mutate()` we can just supply new variables 
after a comma. 
In this case, I add potential outcomes, using `potential_outcomes()`.

```{r}

effect_size_simple <- 70

model_simple <- 
  declare_model(
    rhomis,
    potential_outcomes(Y ~ NFertInput + Z * effect_size_simple)
)

```

The potential outcomes are generated using `potential_outcomes()`.
This will create two variables `Y_Z_1` and `Y_Z_0`, which are 
the potential outcomes if having received treatment (`Z == 1`) or not. 
Note that I will generate `Z` in the next step.

Then it's time to think about assignment. 
I do a clustered randomization, based on the villages:

```{r}

assignment_simple <-
  declare_assignment(Z = cluster_ra(clusters = village, prob = 0.5)) 

```

Next, I declare my theoretical quantity of interest is the treatment 
effect in year 1. 
This is basically my research question.


```{r}

inquiry_simple <-  
  declare_inquiry(ATE = mean(Y_Z_1 - Y_Z_0)) 

```


```{r}

measurement_simple <- declare_measurement(
  Y = reveal_outcomes(Y~Z)
)

```
Since I won't be able to observe Y_Z_1 for the control group, nor Y_Z_0 for the treatment group,
I will need an estimation strategy.
A simple difference in means:

```{r}

estimator_simple <-
  declare_estimator(Y ~ Z,
                    clusters = village, 
                    inquiry = "ATE", 
                    term = "Z", 
                    label = "Simple")

```

Finally, I combine all these elements to declare my design. Note that 
it's only here that R starts actually running the code to randomize 
things. The previous was just declaration!

```{r}
  
design_simple <- model_simple + assignment_simple + inquiry_simple + measurement_simple + estimator_simple
summary(design_simple)

```

Note that the results of my three estimators are identical, which is as expected.


### Diagnosing Design and calculating power

If you want to browse a version of the data created by your design,
use the `draw_data()` function. This is useful to examine the properties of the data.

```{r}

draw_data(design_simple) %>%
  as_tibble()

```

Now to calculate our power. The `diagnose_design()` will run our
model 500 times and our power is simply the fraction of times we
find a statistically significant effect.


```{r}

set.seed(1)
diagnose_design(design_simple)

```

Our power is 0.23, meaning we found a significant result in 23% 
of our model runs.
That's way lower than the traditional 80% threshold.

How large should our effect size be to reach a power of 0.8?
We can use the `redesign()` function for this.
Redesign allows you to vary certain parameters of your design, 
and run the design 500 times for each value of the parameter.
Here I set it to run 200 times, as that takes long enough
already.


```{r diagnose_simple_power, eval = run_diagnosis}

set.seed(1)
diagnosis_simple <-
  design_simple %>%
  redesign(effect_size_simple = seq(100,300,50)) %>%
  diagnose_design(sims = 200) %>%
  tidy() %>%
  as_tibble()

```


```{r, include = FALSE, eval = run_diagnosis}

write_csv(diagnosis_simple,file = here("data/diagnosis_simple.csv"))

```

```{r,  include=FALSE}

diagnosis_simple <- read_csv(file =  here("data/diagnosis_simple.csv"))

```


Now, I want a nice plot. 
For this I filter the tibble I created, and pipe it into `ggplot()`:

```{r}

diagnosis_simple %>%
  filter(diagnosand == "power") %>%
  select(effect_size_simple,power = estimate) %>%
  ggplot(aes(x = effect_size_simple, y = power)) +
    geom_line() +
    geom_point() +
    geom_hline(yintercept=0.8,linetype=2) 

```

We can reliably detect an effect size between 150 and 200 kgs of fertilizer.
The mean fertilizer use is 70kg, so that seems unlikely to be realistic. 
Perhaps we should add covariates to reduce the standard error?

#### Adding more estimators

I define another estimator that includes covariates in
its model specification, and add it to my original design, 
hoping this will increase power.

```{r diagnose_simpe_cov_power, , eval = run_diagnosis}

estimator_simple_cov <-
  declare_estimator(Y ~ Z + LandOwned + educated + female,
                    clusters = village, 
                    inquiry = "ATE", 
                    term = "Z", 
                    label = "Simple Cov")


design_simple_cov <- design_simple + estimator_simple_cov

set.seed(1)
diagnosis_simple_cov <-
  design_simple_cov %>%
  redesign(effect_size_simple = seq(100,300,50)) %>%
  diagnose_design(sims = 200) %>%
  tidy() %>%
  as_tibble()

```



```{r, include = FALSE, eval = run_diagnosis}

write_csv(diagnosis_simple,file = here("data/diagnosis_simple.csv"))

```

```{r,  include=FALSE}

diagnosis_simple <- read_csv(file =  here("data/diagnosis_simple.csv"))

```


```{r}
diagnosis_simple_cov %>%
  filter(diagnosand == "power") %>%
  rename(power = estimate) %>%
  ggplot(aes(x = effect_size_simple, y = power, color = estimator, shape = estimator)) +
    geom_line() +
    geom_point() +
    geom_hline(yintercept=0.8,linetype=2) 

```

Turns out it doesn't help at all in this case.

## Dif in diff

We can of course also opt for a diff-in-diff design. 
But then we have to pretend we have two rounds of data collection.

For that, I prefer to use `dplyr`, 
but it should be possible with `fabricate()` as well.
I first rename all our time-variant varaible to have `_0` at the end, 
create a bunch of correlated varaibles with `_1` at the end, and then `pivot_longer`.


```{r}

effect_size_dd <- 70
sd_dd <- 100

model_dd <- declare_model(
    rhomis %>%
        # add one more year
        rename_with(~ paste0(., "_0"), HHsizeMAE:female) %>%
        mutate(food_secure_1 = correlate(given = food_secure_0, draw_handler = draw_binomial, prob = mean(food_secure_0) , rho = 0.9),
               HHsizeMAE_1 = HHsizeMAE_0 + rnorm(n = nrow(.), mean = 0, sd = 0.1),
               NFertInput_1 = pmax(0,NFertInput_0 + rnorm(n = nrow(.), mean = 0, sd = sd_dd))) %>%
        pivot_longer(ends_with(c("_0","_1")), 
                     names_to = c(".value", "t"),
                     names_pattern = "(.*)_([0-9]+)$",
                     values_drop_na = TRUE,
                     names_transform = list(t = as.integer)),
    potential_outcomes(Y ~ NFertInput + Z * effect_size_dd)


)

```

I modify the inquiry and assignment to only take into account `t == 1`

```{r}

inquiry_dd <-  
  declare_inquiry(ATE = mean(Y_Z_1 - Y_Z_0), subset = (t == 1)) 




assignment_dd <-
  declare_assignment(treatment_group = cluster_ra(clusters = village, prob = 0.5),
                     Z = t * treatment_group)



```

Measurement is unchanged:


```{r}

measurement_dd <- declare_measurement(
  Y = reveal_outcomes(Y~Z)

)

```


Then I use [`feols()`](https://lrberge.github.io/fixest/reference/feols.html)
from the [fixest](https://lrberge.github.io/fixest/index.html)
package to estimate a Two-way fixed effects (2FE) model. 
This can be easily extended to more complex 
designs ([but beware!](https://doi.org/10.1017/pan.2020.33)).

```{r}

library(fixest)
library(broom)

estimator_dd <-
  declare_estimator(Y ~ Z | treatment_group + t,
                    cluster = "village",
                    .method = feols, 
                    term = "Z", 
                    inquiry = "ATE", 
                    label = "2FE")

```

Let's have a look at our power now! 
Of course, in a DiD setting, power depends not only on the effect size, but also on how well the years correlate.
I therefore also vary the `sd_dd` variable, which adds increase between-year variation


```{r diagnose_dd_power, eval = run_diagnosis}

design_dd <- model_dd + assignment_dd + inquiry_dd + measurement_dd + estimator_dd

set.seed(1)
diagnosis_dd <-
  design_dd %>%
  redesign(effect_size_dd = seq(10,90,20),
           sd_dd = c(100,200)) %>%
  diagnose_design(sims = 100) %>%
  tidy() %>%
  as_tibble()

```


```{r, include = FALSE, eval = run_diagnosis}

write_csv(diagnosis_dd,file = here("data/diagnosis_dd.csv"))

```

```{r,  include=FALSE}

diagnosis_dd <- read_csv(file =  here("data/diagnosis_dd.csv"))

```


Our power is now a lot better, even if we turn up the variance. 
However, this all depends on the correlation between years. 
If there is more between-year variation, you lose the benefits of a DiD design.


```{r}

diagnosis_dd %>%
  filter(diagnosand == "power") %>%
  rename(power = estimate) %>%
  mutate(sd_dd = factor(sd_dd)) %>%
  ggplot(aes(x = effect_size_dd, y = power, color = sd_dd, shape = sd_dd)) +
    geom_line() +
    geom_point() +
    geom_hline(yintercept=0.8,linetype=2) 

```

## Propsensity score matching

Finally, let's imagine a Propensity Score matching scenario. We therefore need
to pretend treatment isn't random, but is correlated to the outcome. 

Let's also look at `food_secure`, since it's binary, and thus adds a twist here and there.

Finally, I will use the `resample_data()` function to pretend that we can visit 
more households than are in my original data set.

### Declare Design

Let's put this in declare design.

Treatment assignment is as follows:

- 50% of the villages are selected for treatment. This is just random.
- Roughly 50% of the households are selected. The selection indicator will have
a correlation of `rho` with the household being educated, using `fabricatr`'s 
`correlate()` function. 
Education is also a strong predictor of `food_security`, creating inherent bias.

I compare a standard difference in means, as well as a custom matching function that 
uses the `MatchIt` package to create a balanced data set, that I can then use for
regular regressions, weighted by the propensity score.


```{r}

rhomis_psm <-
  rhomis %>% 
  mutate(Head_EducationLevel = if_else(is.na(Head_EducationLevel),"No_school", Head_EducationLevel),
         income = mean(total_income_USD_PPP_pHH_Yr, na.rm = TRUE) + rnorm(n(), 0, 1000), 
         income = income + if_else(Head_EducationLevel == "primary", 5000, 0) + 
                         if_else(Head_EducationLevel == "secondary", 10000, 0) +
                         if_else(Head_EducationLevel == "adult_education", 20000, 0) +
                         if_else(Head_EducationLevel == "postsecondary", 50000, 0),
         prob_treat = 0.3, 
         prob_treat = prob_treat + if_else(Head_EducationLevel == "primary", 0.1, 0) + 
                         if_else(Head_EducationLevel == "secondary", 0.2, 0) +
                         if_else(Head_EducationLevel == "adult_education", 0.3, 0) +
                         if_else(Head_EducationLevel == "postsecondary", 0.4, 0),         

  ) 


set.seed(1)

library(MatchIt, exclude("select")) # the select function would conflict with dplyr
                                    # this syntax requires R 4.0 or higher

# set parameters
effect <- 10000
#rho <- 0.2


rhomis_psm %>% mutate(sample = draw_binary(prob = prob_treat)) %>% count(sample)
rhomis_psm %>% mutate(sample = simple_ra(N= nrow(.), prob_unit = prob_treat)) %>% count(sample)


population_psm <- declare_population(
  # rhomis %>%
  #   mutate(treated_hh = correlate(given = educated, draw_handler = draw_binary, prob = 0.5 , rho = rho),
  #          potential_outcomes(Y ~ pmin(1,food_secure + Z * draw_binary(effect,nrow(.)))))
  N = nrow(rhomis),
  rhomis,
  #potential_outcomes(Y ~ pmin(1,food_secure + Z * draw_binary(effect,N)))
  potential_outcomes(Y ~ income + Z * effect)
)



assignment_psm <- declare_assignment(
  treated_hh = simple_ra(N= nrow(.), prob_unit = prob_treat),
  treated_vill = cluster_ra(clusters = village, prob = 0.5),
  Z = treated_hh * treated_vill
)
```

I then re-sample from the original data using  `resample_hhs()` I define below.
To include the function in our design, we need to use `declare_step()`

- I ensure I have more control household than treatment households.
- I drop any untreated households in treated villages.
- Compared to the treatment hhs in each cluster, 
I sample twice the number of control hhs.

This simulates that we can't get control households in treated villages, 
for example due to fears of 



```{r}

clusters_treat <- 50
hh_treat <- 15

clusters_control_multi <- 1 # multiplier for oversampling control clusters
hh_control_multi <- 2 # multiplier for oversampling control hhs

resample_hhs <- function(data){
  # this function filters any control households in treated communities
  # and it sets the sample size
  data %>%
    mutate(drop = if_else(treated_vill == 1 & treated_hh == 0,1,0)) %>%
    filter(drop != 1) %>%
      # sample from treated villages and untreated villages
      group_by(treated_vill) %>%
      do({
        if (unique(.$treated_vill) == 1) {
          resample_data(., N = c(village = clusters_treat, hh = hh_treat),unique_labels = TRUE)
        } else {
          resample_data(., N = c(village = clusters_control_multi * clusters_treat, hh = hh_control_multi * hh_treat), unique_labels = TRUE)
        }
      })
}

resampling_psm <- declare_step(
  handler = resample_hhs
)

```

Measurement and inquiry are standard. I also add an OLS estimator to check for bias.

```{r}

measurement_psm <- declare_measurement(Y = reveal_outcomes(Y ~ Z)
) 

inquiry_psm <- declare_inquiry(ATE = mean(Y_Z_1 - Y_Z_0)
)

estimator_notmatched_psm <- 
  declare_estimator(Y ~ Z, 
                    clusters = village_unique,
                    inquiry = "ATE", 
                    term = "Z", 
                    label = "Not matched")


```

Finally, I need to somehow put my matching model in DeclareDesign.
Fortunately, you can put any the output of any function into the `declare_estimator()` function 
using the `handler` argument. 
It does need to accept a `data` argument, 
and output a [`tidy()`](https://broom.tidymodels.org/articles/broom.html) dataframe.

My function `custom_match()` accepts a data frame, uses `MatchIt` to create
a matched data set,
I define a function that does that, and some labelling:

```{r}

custom_match <- function(data, equation, outcome = "Y", term = "Z") {
  
  # the command filter(term == term) wouldn't do anything, so this is a workaround
  term_value <- term

  # run the matching model
  matchit_model <- matchit(equation,
                         data = data,
                         method = "nearest")
  matched_df <- match.data(matchit_model)

  # run a regression, weighted by the propensity score
  lm_robust(Y ~ Z, clusters = village_unique, data = matched_df, weights = weights) %>%
    tidy() %>%
    mutate(outcome = outcome) %>%
    filter(term == term_value)
}

estimator_matched_psm <- declare_estimator(
  Z ~ HHsizeMAE + educated + LandOwned + NFertInput + female + NFertInput,
  handler = label_estimator(custom_match),
  inquiry = "ATE",  # Tie the estimator to the declared inquiry
  label = "Matched",
  term = "Z",
)

design_psm <-
  population_psm +  inquiry_psm + assignment_psm + resampling_psm + measurement_psm + estimator_notmatched_psm + estimator_matched_psm

```


### Checking balance

It's useful to explore the property's of our matching estimator. 
First  I draw data based on my design and create matched version of that data.

```{r}

set.seed(1)

test_data <- 
  draw_data(design_psm) %>%
  as_tibble() 


  matchit_model <- matchit(Z ~ HHsizeMAE + educated + LandOwned + NFertInput + female + NFertInput,
                         data = test_data,
                         method = "nearest")
  matched_df <- match.data(matchit_model)

```

Then I create balance table for the `test_data`:

```{r}

library(modelsummary)
library(flextable)


# balance table
test_data %>% 
  mutate(Z = factor(Z, labels = c("Control", "Treatment"))) %>%
  select(Z, Y, food_secure, HHsizeMAE, educated, LandOwned, NFertInput, female, NFertInput) %>%
  datasummary_balance( ~ Z , data = ., 
                      output = "flextable", stars = TRUE, 
                      dinm = TRUE, dinm_statistic = "p.value") %>%
  fix_border_issues() %>% 
  autofit()

```

And for the `matched_df`, which is clearly more balanced.
Note that according to the [documentation](https://modelsummary.com/reference/datasummary_balance.html#global-options),
`datasummary_balance()` reports weighted means if a `weights` variable is present. 

```{r}

matched_df %>% 
  mutate(Z = factor(Z, labels = c("Control", "Treatment"))) %>%
  select(Z, Y, food_secure, HHsizeMAE, educated, LandOwned, NFertInput, female, NFertInput) %>%
  datasummary_balance( ~ Z , data = ., 
                      output = "flextable", stars = TRUE, 
                      dinm = TRUE, dinm_statistic = "p.value") %>%
  fix_border_issues() %>% 
  autofit()

```


Let's run the models two times:


```{r}

unmatched_model <- lm_robust(Y ~ Z, clusters = village_unique, data = test_data)
matched_model <- lm_robust(Y~Z, clusters = village_unique, data = matched_df, weights = weights)

modelsummary(list(lm = unmatched_model, psm = matched_model), output = "flextable",
             gof_map = c("nobs","r.squared","adj.r.squared"))

```

Both of these underestimate the treatment effect by a bit, as it should be around 0.2.

### Bias

Now, let's diagnose the design!

First, let's examine bias, given selection effects. 
The strength of the selection effect is given by `rho`,
so I use `redesign()` to vary it:

```{r diagnose_psm_bias, eval = run_diagnosis}

diagnosis_psm_bias <-
  design_psm %>%
  redesign(rho = seq(from = 0, to = 0.9, by = 0.3))  %>%
  diagnose_design(sims = 100) %>%
  tidy() %>%
  as_tibble()

```

```{r, include = FALSE, eval = run_diagnosis}

write_csv(diagnosis_psm_bias,file = here("data/diagnosis_psm_bias.csv"))

```

```{r,  include=FALSE}

diagnosis_psm_bias <- read_csv(file =  here("data/diagnosis_psm_bias.csv"))

```

And I put the result in a plot:

```{r}

diagnosis_psm_bias %>%
  filter(diagnosand == "bias") %>%
  select(rho, bias = estimate, estimator) %>%
  ggplot(aes(x = rho, y = bias, 
             shape = estimator, color=estimator)) +
    geom_line() +
    geom_point() 

```

Clearly, the more treatment is correlated to education, 
the higher the bias in the regular model.
Bias is never extreme, however.


### Power 

Then, let's examine power. 


```{r diagnose_psm_power, eval = run_diagnosis}

diagnosis_psm_power <-
  design_psm %>%
  redesign(clusters_treat = c(20,40,60),
           effect = seq(0.1,0.4,0.1)) %>%
  diagnose_design(sims = 100) %>%
  tidy() %>%
  as_tibble()

```

```{r, include = FALSE, eval = run_diagnosis}

write_csv(diagnosis_psm_power,file = here("data/diagnosis_psm_power.csv"))

```

```{r,  include=FALSE, message = FALSE}

diagnosis_psm_power <- read_csv(file =  here("data/diagnosis_psm_power.csv"))

```

When we plot the results, it seems sample size doesn't affect power much 
(not sure why, perhaps the resampling?). 
The design can pick up an `effect` of 0.4,
meaning that 40% of the food insecure beneficaries become food secure.
The actual treatment effect would be lower, as some beneficiaries would
have been food secure before the intervention.


```{r}

diagnosis_psm_power %>%
  filter(diagnosand == "power") %>%
  select(clusters_treat, power = estimate, estimator, effect) %>%
  ggplot(aes(x = clusters_treat, y = power, 
             shape = estimator, color=estimator)) +
    geom_line() +
    geom_point() +
    geom_hline(yintercept=0.8,linetype=2) +
    facet_wrap(~effect)

```

