# Functions {#functions}

Functions are great to re-use and structure your code.

## Basic structure

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}

add_two <- function(argument) {
  # Do something with the input
  return <- argument + 2

  #call the intended output in the last line
  return
}

add_two(84)

```

## Using functions in a pipe

By default, a function in a pipe uses the dataset of the pipe as its first argument.
Also make sure to include a quick description of what your function does.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}
library(tidyverse)

add_number_column <- function(data,number=2) {
  #This functions takes a data frame, and returns the same data frame,
  #but with an exta column of 2sas its first variable. 
  #The user can specify a number other than 2 using the num argument
  #This is great if you want a column of twos as the first variable
  #of your data set.
  data %>%
    mutate(bestvar = number) %>%
    select(bestvar,everything())
}

read_csv(here("data/SAFI_clean.csv"),na = "NULL") %>%
add_number_column() 

read_csv(here("data/SAFI_clean.csv"),na = "NULL") %>%
add_number_column(4) 

```

## Passing variables to a function

The following doesn't work, because R can't find `no_membrs`, even though it 
does exist in the data set in the pipe.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=TRUE}

add_two_to_a_column_broken <- function(data,variable) {
  #function that wants to add two two a column specified by the user
  #but it doesn't work.
  data %>%
  mutate(variable = variable + 2)
}

read_csv(here("data/SAFI_clean.csv"),na = "NULL") %>%
add_two_to_a_column_broken(no_membrs) 

```

For a technical note on why this is, 
see [here](https://dplyr.tidyverse.org/articles/programming.html).
To solve this, wrap the variable in your function definition
in `{{` and `}}`, and change the `=` in `mutatate` to `:=`. 


```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

add_two_to_a_column_fixed <- function(data,variable) {
  #function that adds two two a column specified by the user
  data %>%
  mutate( {{variable}} := {{variable}} + 2)
}

read_csv(here("data/SAFI_clean.csv"),na = "NULL") %>%
  add_two_to_a_column_fixed(no_membrs)

```

## Function to create dummies from text

Remember our code to generate dummies based on text:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}

read_csv("data/SAFI_clean.csv", na = "NULL") %>%
    mutate(items_owned = ifelse(is.na(items_owned),"None",items_owned)) %>%
    select(key_ID,items_owned) %>%
    separate_longer_delim(items_owned, delim = ";") %>%
    mutate(value = 1) %>%
    pivot_wider(names_from = items_owned,
                values_from = value,
                names_prefix = "owns_",
                values_fill = 0) 

```

I now want to apply this same principle to `months_lack_food` column.
I could copy-paste the code, but the better solution is to make a function:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}


create_dummies <- function(df,stringvar,prefix="",delim = ";") {
  #Function that generates dummies from a string variable
  #containing multi-response answers
  #returns dataset without original var, but with dummies.

  df %>%
    separate_longer_delim({{stringvar}}, delim = delim) %>%
      mutate(value = 1) %>%
        pivot_wider(names_from = {{stringvar}},
                    values_from = value,
                    names_prefix = prefix,
                    values_fill = 0) 
}

read_csv(here("data/SAFI_clean.csv"), na = "NULL") %>%
  mutate(items_owned = ifelse(is.na(items_owned),"None",items_owned)) %>%
  create_dummies(stringvar = items_owned, prefix = "owns_") %>%
  create_dummies(months_lack_food,"lack_food_")

```

The code is now clear, I can update the function in one place and all variables will
be updated accordingly. Plus, if I want to use the same code in another project, I can
easily do it. 

## Using a function factory to create multiple variables

A function can also return a function. This means that you can use `map()` to
make a list of functions. This can be handy for `mutate(across()`, which can
take a list of function to create multiple variables. Using this to solve
the problem of splitting strings:


```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}


items <- c("bicycle","television","solar_panel","table","cow_cart","radio",
           "cow_plough","solar_torch","mobile_phone","motor_cycle","fridge",
           "electricity","sofa_set","lorry","sterio","computer","car")


# define the function factory first: all it does is return another function, 
# that can only extract one item
make_splitfunction <- function (item) {
  function(var){
    str_count(var,item)
  }
}

#make a list with one function for each item
function_list <-
  items %>%
  set_names(items) %>%
  map(make_splitfunction)


#apply the list of functions using mutate(across)
read_csv(here("data/SAFI_clean.csv"), na = "NULL") %>%
  mutate(across(.cols = items_owned,
                .fns = function_list,
                .names = "owns_{.fn}")) %>%
  select(items_owned,starts_with("owns_"))


```

Or, if you want fewer (but more complex) steps, you can generate the function
list on the fly within `mutate(across())`.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

read_csv(here("data/SAFI_clean.csv"), na = "NULL") %>%
  mutate(across(.cols = items_owned,
                .fns = 
                  items %>%
                  set_names(items) %>%
                  map(function(item) function(var) str_count(var,item)),
                .names = "owns_{.fn}")) %>%
  select(items_owned,starts_with("owns_"))

```