# Data Exploration

The first step of using data is exploring it. I will use Stata data
because it has labels, making it easy to get a sense of the data once
you get the hang of how labels are dealt with in R.

## Setup

Download the [cars data set from here](http://www.principlesofeconometrics.com/stata.htm)
or run the code below:

```{r eval=FALSE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}


download.file(
  "https://raw.githubusercontent.com/kleuveld/r_cheatsheet/main/data/cars.dta",
  here("data/cars.dta"), mode = "wb"
)

```

## Making a codebook from a Stata .dta

In Stata variables have labels, which is great because they're more informative
than variable names. In R, it can be a bit tricky to access the labels of imported
dta's, but making a code book isn't that hard.


First, load the cars data set:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

library(tidyverse)
library(haven)
library(here)


cars <- read_dta(here("data/cars.dta"))

```

The variable labels are stored as attributes of the variables. The `attributes()`
function returns all attributes:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

attributes(cars$mpg)

```

 To see only the label use:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

attributes(cars$mpg)$label

```

To create a data frame with all variable labels we can apply `attributes()`
to all variables using `map_chr()` from the `purrr` package: 

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

codebook <- 
  tibble(var = colnames(cars),
         label = map_chr(cars,~attributes(.x)$label)) 

codebook

```

To make it slightly more useful, we can add some summary statistics. 
I can apply a number of functions to a data frame using `map_dbl()`, 
which returns a named vector:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

list_of_functions <- list(mean=mean,sd=sd,min=min,max=max)

list_of_functions %>%
  map_dbl(~.x(cars$mpg, na.rm = TRUE))


```

To do this for every column in a dataframe, I wrap the code above 
in a function, and use `map()` to apply that function to the columns. 


```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

stats_to_tibble <- function(var,funs) {
  funs %>%
    map_dbl(~ifelse(is.numeric(var),.x(var,na.rm = TRUE),NA)) %>%
    as_tibble_row()
}

summ_stats <- 
  cars %>%
    map(~stats_to_tibble(.x,list_of_functions)) %>%
    list_rbind()
summ_stats



```

I can bind that with the codebook I had before to get a nice 
overview of all the variables in my dataset:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

bind_cols(codebook, summ_stats)   

```

Here's a re-usable function that add more columns, handles empty labels
(using `coalesce()`) and rounds the output so it's human-readable:


```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

create_codebook <- function(.df,stats = list(mean=mean,sd=sd,min=min,max=max,
                                            prop_miss=prop_miss)) {
  labels <- tibble(var = colnames(.df),
                   label = map_chr(.df,function(x) coalesce(attributes(x)$label,"")),
                   type = map_chr(.df, typeof))


  prop_miss <- function(x,na.rm = TRUE) {
    mean(is.na(x))
  }

  stats_to_tibble <- function(var,stats) {
    map_dbl(stats,~ifelse(is.numeric(var),.x(var,na.rm = TRUE),NA)) %>%
    as_tibble_row()
  }

  sumstats <-
    .df %>%
    map(~stats_to_tibble(.x,stats)) %>%
    list_rbind() %>%
    mutate(across(where(is.numeric),
                ~round(.x,2)))  

   bind_cols(labels,sumstats)
}

create_codebook(cars) 


```

## Correlogram

Another great data exploration tool is the correlogram, which displays
the correlations between many variables. To create one, I use `ggpairs()`
from the `GGally` package:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

library(GGally)

ggpairs(cars) 

```

You can also split the correlogram by a variable, like I do with the number
of cylinders below:


```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

cars %>%
  ggpairs(columns = c(1,3,4), 
          ggplot2::aes(colour=factor(cyl))) 

```