# Programming with R {#programming}

You can improve your R code by using programming features such as 
loops and functions to make your code reusable. This prevents errors, 
while saving you time.


## Loops

Loops repeat code a number of time, they have the following structure:


```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

for (i in 1:4){
	print(i)
}

```

The `i` is the iterator, and each iteration of the loop it takes on the value of
a different element of the object that follows the `in`; in this case, `i` takes 
on 1,2,3 and finally 4 as its values. Therefore, this code is equivalent to:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

print(1)
print(2)
print(3)
print(4)

```

But it's much easier to maintain! Imagine having to do 86 iterations: you would
need to copy paste the code 85 times every time you want to update something. 
Using loops, you don't need to copy-paste a thing, leading to fewer errors.

While loops can be a perfectly fine way of  doing repeated work, most R 
programmers prefer not to use loops, but instead use functions such as 
`lapply()` and `map()`. However, for this you will need to know functions.

## Functions {#functions}

### Basic Functions

A function is an R object that's essentially a shortcut to run a bit of code. 
Usually, a function takes an argument (the input), and returns an output. The 
argument is always provided within parentheses. The  `sqrt()` function, for 
example, takes a number as its argument and returns the square root.

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

sqrt(4)

```

The beauty of functions is that they allow you to define a set of operations in 
one place, and re-use it as often as you want. If you then need to update the
operations, all you need to do is update the function, instead of hunting down 
every instance of the operations throughout your script.

Take for example the following code that cleans missing data.


```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

#define an example variable
variable1 = c(1,2,3,-99,5,6)


#change missing
variable1_cleaned = if_else(variable1 == -99, NA,variable1)

variable1_cleaned 


```

This works, but I need to do this at many different places in my code, and when
copying the code to other places, it's easy to make small mistakes. Moreover, if
I notice my procedure for cleaning missing data is wrong, I will have to hunt 
down all the places I've copied this code to, and change it (which, again, is 
an error prone process). 

A common piece of advice is to put any code that you
have to copy-paste more than twice in a function. Once I do that, I can just 
call  the function wherever I need to clean data; if I want to change how I deal 
with missings, I now only have to make changes in one place:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

clean_missing <- function(input_data) {

	returned_data <- if_else(input_data == -99, NA, input_data)
	
	return(returned_data)
}


```

I've pretty much copied the code I had above. Note that I could have used any 
name for my argument (defined in the parentheses after function); I just thought 
`input_data` made sense. I use the `return()` function to return data.

I can now just use the `clean_data()` wherever I want:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

variable1_cleaned <- clean_missing(variable1)

variable1_cleaned

```

Note that any variables we created in the function are kept within the function, 
and you can't access them later:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

variable1_clean <- clean_missing(variable1)
returned_data


```

This keeps your working environment nice and clean, which again prevents 
problems.

Above I used the return function. This is not needed. Anything on the 
last line of the function that would normally be returned to the console, is
used a return value of the function:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

clean_missing <- function(input_data) {

	if_else(input_data == -99, NA, input_data)
	
}

clean_missing(variable1)


```

### Functions and the tidyverse

In the previous function I passed a vector to clean, what if want to use it
in a data cleaning pipe? 

It's useful to start with the code without a function, and then to generalize
from there:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}


data_frame <- tibble(variable1 = variable1)

data_frame %>%
	mutate(variable1 = if_else(variable1 == -99,NA,variable1))

```

Let's try to put that in a function. The function will need to take 
a data frame and a variable name as its arguments. Results from one step in the 
pipe, are passed as the first argument for the next step. So make sure the data 
is the first argument! 

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

clean_missing_df <- function(input_dataframe, variablename){
	input_dataframe %>%
		mutate(variablename = if_else(variablename == -99,NA, variablename))
	
}

data_frame %>%
	clean_missing_df(variable1)


```

Wait, something went wrong! R just created a new variable called "variablename".
For a technical note on why this is, see 
[here](https://dplyr.tidyverse.org/articles/programming.html). In short, we want
R too look what's *in* `variablename`, not just use it as is. To do so, wrap the 
variable name in your function definition
in `{{` and `}}`, and change the `=` in `mutatate` to `:=`. 

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}


clean_missing_df <- function(input_dataframe, variablename){
		input_dataframe %>%
		mutate({{ variablename }} := if_else({{ variablename }} == -99,
											 NA,
											 {{ variablename }}))
	
}


data_frame %>%
	clean_missing_df(variable1)


```

### Example 1: Re-using plot specifications {#plot-function}

If you're writing a report that includes the same type of plot in many
places, it makes sense to make it into a function. 

Let's take the [stacked bar plot from the plotting chapter](#stacked-wur):

```{r}

library(ggthemewur)

stacked_bar_plot <- function(df, by, fill) {
  # outputs a stacked bar chart
  
  df %>%
    mutate( {{ by }} := count_label( {{ by }})) %>%

    ggplot(aes(y = {{ by }}, fill = {{ fill }})) +
      geom_bar(position = position_fill(reverse = TRUE)) +
      labs(x= "") +
      theme_wur() +
      scale_fill_wur_discrete() +
       geom_text(stat = "count", 
                 aes(label = after_stat(count)), 
                 position = position_fill(vjust = 0.5, reverse = TRUE), 
                 color = "white") +
      scale_x_continuous(labels = scales::percent_format()) +
      theme(legend.position="bottom") +
      guides(fill = guide_legend(nrow = 1, 
                                 title.position="top", 
                                 title.hjust = 0.5)) 
}

count_label <- function(vector) {
  # takes a vector of strings or factor, 
  # output a factor vector with N = N included in the labels
  fct_recode(factor(vector), 
             !!!vector %>%
               as_tibble %>%
               group_by(value) %>%
               summarize(n = n()) %>%
               mutate(value = as.character(value),
                      newlabel = paste0(value,"\nn=",n)) %>%
               pull(value, name = newlabel))
  
}


read_csv(here("data/SAFI_clean.csv"), na = "NULL") %>%
  stacked_bar_plot(village,respondent_wall_type)


read_csv(here("data/SAFI_clean.csv"), na = "NULL") %>%
  stacked_bar_plot(village,affect_conflicts)

```


### Example 2: Function to create dummies from text

Remember our code to generate dummies based on text:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}

read_csv("data/SAFI_clean.csv", na = "NULL") %>%
    mutate(items_owned = ifelse(is.na(items_owned),"None",items_owned)) %>%
    select(key_ID,items_owned) %>%
    separate_longer_delim(items_owned, delim = ";") %>%
    mutate(value = 1) %>%
    pivot_wider(names_from = items_owned,
                values_from = value,
                names_prefix = "owns_",
                values_fill = 0) 

```

I now want to apply this same principle to `months_lack_food` column.
I could copy-paste the code, but the better solution is to make a function:

```{r eval=TRUE,echo=TRUE,warning=TRUE,error=TRUE,message=TRUE}


create_dummies <- function(df,stringvar,prefix="",delim = ";") {
	#Function that generates dummies from a string variable
	#containing multi-response answers
	#returns dataset without original var, but with dummies.

	df %>%
		separate_longer_delim({{ stringvar }}, delim = delim) %>%
	    mutate(value = 1) %>%
	    pivot_wider(names_from = {{stringvar}},
	                values_from = value,
	                names_prefix = prefix,
	                values_fill = 0) 
}

read_csv(here("data/SAFI_clean.csv"), na = "NULL") %>%
	mutate(items_owned = ifelse(is.na(items_owned),"None",items_owned)) %>%
	create_dummies(stringvar = items_owned, prefix = "owns_") %>%
	create_dummies(months_lack_food,"lack_food_")

```

The code is now clear, I can update the function in one place and all variables will
be updated accordingly. Plus, if I want to use the same code in another project, I can
easily do it. 


## Map()

`map()` works just like loop, but generally faster. It takes two arguments:

- An iterable object, like a vector or list
- A function

It will then apply the function to each element of the object. It will return
a list with the results of each iteration.

For example, to get the the square root for each of the numbers in a vector:

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

vector <- c(4,9,16) 
map(vector,sqrt)

```

You may not quite like the fact that map returns a list, but quite a few things in
R use lists. `modelsummary()` for example. Suppose I want to regress the same y on
a number of combinations of independent variables, I can put the combination of
varaibles in a list, and then run my model for each element of the list.

I use `reformulate()` to convert strings into a formula 'lm()' can use. Also,
I use `~` to create 
[a purrr-style inline anonymous function](https://adv-r.hadley.nz/functionals.html#purrr-shortcuts)
within `map()`. 

```{r eval=TRUE, include = TRUE, echo=TRUE, warning=TRUE, error=TRUE, message=FALSE}

library(modelsummary)

# generate some sample data
df <- tibble(y = rnorm(50),x1 = rnorm(50),x2 = rnorm(50))


# define a list of models: 
# the left-hand sides are the labels
# the right-hand sides the independent variables I will pass to reformulate.

models <-
list("X1 only" = "x1", 
	 "X2 only" = "x2", 
	  "Both" = c("x1","x2"))


# map() works nicely in  dplyr pipe!
# the ~ creates a purrr-style anonymous function, where .x is current element of the list
models %>%
	map(~lm(reformulate(.x, response="y"),data = df)) %>%
	modelsummary(output = "flextable")


```


